Universal Anonymous Signatures

### 2022-08-17 prism-backend/research/gsac/introduction-paper.tex
Applying changes after pass.

@@

%% Brief description of main precedents (simple words)
\paragraph{Anonymous authentication and signatures.} %
Privacy-preserving digital signature and authentication systems have been a
prolific field of study in the last few decades. Anonymous credentials (AC)
\cite{chau85} and group signatures (GS) \cite{ch91} are among the most relevant
schemes. While there are many variants, the high-level idea is common: after
obtaining a credential, users can authenticate anonymously and indistinguishably
from other users with similar credentials.

@@

\paragraph{Utility at authentication time and after authentication time.} %
AC schemes focus on utility at authentication time. For instance, encoding
multiple attributes within a credential and allowing to selectively disclose a
subset of them, or even arbitrary predicates. The real world applications of
such systems are direct, e.g., to prove ``Over 18''-type of claims. Indeed, the
scientific research behind anonymous credentials is making its own way to become
a foundational block of the technology some call ``anoncreds''%
\footnote{\url{https://hyperledger-indy.readthedocs.io/projects/sdk/en/latest/docs/design/002-anoncreds/README.html}. Last access, June 18th, 2022.},
the privacy-preserving cousin of W3C's Verifiable Credentials%
\footnote{\url{https://www.w3.org/TR/vc-data-model/}. Last access, June 18th,
  2022.}.

@@

%% Main drawbacks of precedents
\paragraph{Limitations of GS and AC.} %
However, even the generalisation efforts that have independently taken place in
GS and AC, are quickly seen to have limited applicability in the real world.
On the GS side, even though MPS achieves a nice balance on when a signature is
deanonymizable and what information can be extracted from it, upon closer look,
this information is somehow meaningless. Namely, the notion of identity is not
well defined and, thus, it is not possible to claim whether the value output by
their generalised open function makes sense or even if it is legitimate. Also,
just being able to prove membership at signing time is typically not enough, and
one needs to prove more general claims. That is precisely what AC schemes do,
and AC schemes also have a well defined notion of identity via the credential
attributes. Precisely, it is this defined concept of identity what enables AC
schemes to state when some information revealed at authentication time is a
forgery or not. On the other hand, the notion of accountability  is barely
present in AC schemes. For instance, non-frameability is not even considered
-- and it is non-frameability what ensures when a user cannot be held
accountable.
%
In any case, perhaps the most impractical limitation of all is that too many
models exist in both fields. Although similar, each model focuses on slightly
different utility variants, making them hardly comparable. And, of course, AC
focus only on utility at authentication (signing) time, while GS focus on
utility after signing (authentication) time. To the best of our knowledge, no
model exists that combines both, in a general enough manner. In practice, this
forces system designers to decide which one they do not want to lose -- or,
which one they will lose.

@@

\paragraph{General model for anonymous signatures.} %
Acknowledging that the strengths of AC are the weaknesses of GS, and vice versa,
we come up with a unified model. In a nutshell, we give a syntax and model for
anonymous signatures -- which can be turned into an interactive authentication
system -- with attribute-enabled credentials, and generalised utility both at
authentication/signature-time, and after. Concretely, we propose a scheme
through which it is possible to model revealing arbitrary claims at signing time
(as in AC), enable conditional release of utility information after signing time
(as in MPS) and, more importantly, come up with the combination of the previous
that is more suitable for each scenario. We also generalise the credential
issuance process, subsuming an even larger set of previously known schemes.
Given its generalisation power, we refer to schemes meeting this model as
Universal Anonymous Signatures.

\paragraph{Generic construction, and connections with previous schemes.} %
We give a generic construction based on well-known building blocks, and prove
its security. To prove its universality, we show how restrictions of our generic
construction directly imply the direct ancestors of UAS: group signatures and
anonymous credentials. This includes not only their vanilla versions, but also
relevant variants, such as MPS \cite{ngsy22}, (a modified version of) group
signatures with message-dependent opening \cite{seh+12}, or anonymous
credentials with revocation \cite{cks10}. We also show how minimal variations in
our generic construction allow us to build ring signatures \cite{rst06} or
delegatable anonymous credentials \cite{bcc+09}. The most interesting part being
that, while we prove implications with other previously known schemes, our
generalisation approach allows to merge features of them: e.g., one can easily
model and instantiate GSs with with no revocation at all, ACs with full
de-anonymization, or anything in between -- like being able to deanonymize only
if certain conditions are met.

%% Our results; Tech nuances
\paragraph{Modelling challenges.} %
Combining and extending features of GS and AC into a unified model unavoidably
brings complexity. For instance, we allow proving arbitrary claims over
previously obtained credentials at issuance time, in order to inform the
issuance decision. Thus, it is necessary to model unforgeability guarantees
already at that point -- something not considered in previous works. Also, in
order to meet the general practice in AC, the model for \UAS includes multiple
issuers, requiring careful bookkeeping to prevent trivial wins. Interestingly,
combining this generalisation with the GS world also paves the way to
include multiple openers, each with its own opening functionality -- also
requiring careful bookkeeping. While powerful from a practical point of view,
this introduces similar limitations privacy-wise as in functional encryption
schemes \cite{bsw11} -- e.g., revealing too granular information at signing or
opening time makes privacy go away. However, on this occasion, it is the signer
(perhaps jointly with the verifier) who decides the signing and
opening policy. Thus, the user is in full control of how deanonymizable its
signatures will be, and knows it at signature generation time. This flexibility
on the opening values also leads to the
need of using extraction-based techniques for the modelling, since we cannot
expect that opening outputs the real identity of the signer. This comes as no
surprise, though, as it was already observed in GS schemes with non-conventional
opening \cite{dl21}, in DAA schemes \cite{cdl16,cdl16b} or in AC schemes that
need to ensure correctness of information generated before authentication, like
delegatable credentials \cite{bcc+09}, that prove validity of the delegation
chain. In fact, simulation-extractability seems to be most useful here,
something that was also observed in \cite{cl06} for producing signatures out of
knowledge of witnesses for concrete NP statements.

@@

\paragraph{Sample real-world use cases.} %
By enabling customization at issuance, signing and opening times, the
possibilities are large. As an example, consider a social network that requires
under-age users to get permission from one parent. This can be defined via a
delegatable credential scheme, checking that the delegation chain of the user
requesting to join is valid via a policy at issuance time. Later, in order to
make some post in the system advertising something to be sold in a local
second-hand page in the social network, the user may need to reveal, at signing
time, being a citizen of the local area (e.g. city). This can be done via an
appropriate policy at signing time. If after the product is sold, it proves
to be defective, the social network may directly output the email of the user
if s/he is over-age, or the email of the adult in charge if s/he is under-age.
And that could be achieved via a policy for utility after signing time (at
opening time). Finally, if, in the same social network, some other
issuer-user-verifier situation comes up that requires a different set of utility
choices, just the appropriate functions would need to be changed -- and, given
that our model allows multiple issuers and openers, both could coexist.

%% Our techniques
%%% E.g., describe how we compatibilize both with AC sel disc., ring sigs, and MPS
\paragraph{Our construction approach.} %
The high-level idea for our generic construction \CUASGen is simple. Credentials
are signatures over blocks of messages (some signed in committed form), where
each message can be a different attribute, and a user who wants to
simultaneously leverage multiple credentials needs to bind all of them to the
same key. We add three generic functions: \fissue, leveraged by users at
credential issuance time to prove claims over attributes in other credentials,
in order to get a new one; \feval, proving correctness of the utility revealed
at signing time; and \finsp, proving correctness of the utility information
extractable after signing. The latter is, in addition, fed and modulated with
the output of \feval. The output of \feval is
actually double: one part is shared in the clear along with the signature, and
the other is encrypted, so that it can be used to tweak -- in a privacy-enhanced
manner -- the behaviour of \finsp. The output of \finsp is encrypted with an
opener's key. The pattern of the statements proven in each function is
summarised in \figref{fig:proof-blocks-uas}.

@@

From \CUASGen, one can already build known (and more
functionality-limited) schemes from what we call \emph{restrictions} of
\CUASGen simply by giving concrete definitions of \fissue, \feval, and \finsp.
Yet, the model remains the same, and we prove that if such restrictions of
\CUASGen are secure in the \UAS model, they lead to secure constructions for
the related primitives (GS variants, AC variants, etc). To define our generic
construction, we also specify concrete NP relations for the underlying NIZK
proofs for issuing, signing, and opening. These NP relations must not be
considered as immutable, though. Actually, we show how, by making a minor
variation into the one used for signing, a \CUASGen restriction can securely
build ring signatures.

We emphasize that, while -- already before \UAS -- it may have not seemed hard
to build either of these primitives from another, the core of our contribution
is that, with \UAS, one just needs to give concrete definitions of three
functions (or little more). Security of the final primitive then essentially
follows from the fact that it is a secure instantiation of an \UAS scheme. We
believe that this is a powerful notion, that may finally allow research lines
(AC and GS) that have followed different paths to benefit from the findings of
each other. \figref{fig:relations} graphically summarises the relations we prove
between \UAS and other schemes -- although, of course, many more are probably
achievable.

\begin{figure}[ht!]
  \centering
  \scalebox{0.9}{
    \input{figures/relations.tex}
  }
  \caption{Summary of schemes that can be instantiated as restrictions of our
    generic constructions, $\CUASGen^{\ast}$, and proved secure under our \UAS
    model. Between parenthesis we show the adopted reference model, and the
    section offering full detail.}
  \label{fig:relations}
\end{figure}

>>>>

\todo{Mention Camenisch, Krenn and Lehmann's work.}

@@

%% Brief description of main precedents (simple words)
\paragraph{Anonymous authentication and signatures.} %
Privacy-preserving digital signature and authentication systems have been a
prolific field in the last few decades. Anonymous credentials (AC)
\cite{chau85} and group signatures (GS) \cite{ch91} are among the most relevant
schemes. For both, the high-level idea is common: after obtaining a credential,
users can authenticate anonymously and indistinguishably from other users with
similar credentials.

@@

\paragraph{Utility at authentication time and after authentication time.} %
AC schemes focus on utility at authentication time. For instance, encoding
multiple attributes within a credential and allowing to selectively disclose a
subset of them. The real world applications of such systems are direct, e.g.,
to prove ``Over 18''-type of claims. Indeed, the scientific research behind
anonymous credentials is making its own way to become a foundational block of
the technology some call ``anoncreds''%
\footnote{\url{https://hyperledger-indy.readthedocs.io/projects/sdk/en/latest/docs/design/002-anoncreds/README.html}. Last access, June 18th, 2022.},
the privacy-preserving cousin of W3C's Verifiable Credentials%
\footnote{\url{https://www.w3.org/TR/vc-data-model/}. Last access, June 18th,
  2022.}.

@@

%% Main drawbacks of precedents
\paragraph{Limitations of GS and AC.} %
However, even the generalisation efforts that have independently taken place in
GS and AC, are quickly seen to have limited applicability in the real world. On
the GS side, even though MPS achieves a powerful balance on when a signature is
deanonymizable and what information can be extracted from it, upon closer look,
this information is somehow meaningless. Namely, the notion of identity is not
well defined and, thus, it is not possible to claim whether the value output by
their generalised open function makes sense or even if it is legitimate. Also,
just being able to prove membership at signing time is typically not enough, and
one needs to prove more general claims. That is precisely what AC schemes do,
while also providing a well defined notion of identity via credential
attributes. Precisely, it is this defined concept of identity what enables AC
schemes to state when some information revealed at authentication time is a
forgery or not. On the other hand, the notion of accountability  is barely
present in AC schemes. For instance, non-frameability is not even considered
-- and it is non-frameability what ensures when a user cannot be held
accountable.
%
In any case, perhaps the most impractical limitation of all is that too many
models exist in both fields. Although similar, each model focuses on slightly
different utility variants, making them hardly comparable. And, of course, AC
focus on utility at authentication (signing) time, while GS focus on
utility after signing (authentication) time. To the best of our knowledge, no
model exists that combines both, in a general enough manner. In practice, this
forces system designers to decide which one they do not want to lose -- or,
which one they will lose.

@@

\paragraph{General model for anonymous signatures.} %
Acknowledging that the strengths of AC are the weaknesses of GS, and vice versa,
we come up with a unified model. In a nutshell, we give a syntax and model for
anonymous signatures -- which can be turned into an interactive authentication
system -- with attribute-enabled credentials, and generalised utility both at
signature time, and after. Concretely, we propose a scheme
through which it is possible to model revealing arbitrary claims at signing time
(as in AC), enable conditional release of utility information after signing time
(as in MPS) and, more importantly, come up with the combination of the previous
that is more suitable for each scenario. We also generalise the credential
issuance process, subsuming an even larger set of previously known schemes.
Given its generalisation power, we refer to schemes meeting this model as
Universal Anonymous Signatures.

\paragraph{Generic construction, and connections with previous schemes.} %
We give a main generic construction based on well-known building blocks, and
prove its security. To prove its universality, we show how restrictions of our
generic construction directly imply the direct ancestors of UAS: group
signatures and anonymous credentials. This includes not only their vanilla
versions, but also relevant variants, such as MPS \cite{ngsy22}, (a modified
version of) group signatures with message-dependent opening \cite{seh+12}, or
anonymous credentials with revocation \cite{cks10} or delegation \cite{bcc+09}.
Strictly speaking, for AC schemes, we build on a simple transform of our main
construction and model that turns non-interactive signatures into interactive
authentication protocols with protection against replay attacks\footnote{This is
  the well-known folklore approach that has the verifier send a fresh nonce to
  be signed.}. We further show how to build ring signatures (RS) in the model
given by \cite{bkm06}, from an alternative construction that improves issuer
privacy.
%
The most interesting part being that, while we prove implications with other
previously known schemes, our generalisation approach allows to merge features
of them: e.g., one can easily model and instantiate GSs with with no revocation
at all, ACs with full de-anonymization, RS with linkability, or anything in
between -- like being able to deanonymize only if certain conditions are met.

%% Our results; Tech nuances
\paragraph{Modelling challenges.} %
Combining and extending features of GS and AC into a unified model unavoidably
brings complexity. For instance, we allow proving arbitrary claims over
previously obtained credentials at issuance time, in order to inform the
issuance decision. Thus, it is necessary to model unforgeability guarantees
already at that point -- something not considered in previous works. Also, in
order to meet the general practice in AC, the model for \UAS includes multiple
issuers. Interestingly, combining this generalisation with the GS world also
paves the way to include multiple openers, each with its own opening
functionality. While powerful from a practical point of view, this introduces
similar (unavoidable?) limitations privacy-wise as in functional encryption
schemes \cite{bsw11} -- e.g., revealing too granular information at signing or
opening time makes privacy go away. However, on this occasion, it is the signer
(perhaps jointly with the verifier) who decides the signing and
opening policy. Thus, the user is in full control of how deanonymizable its
signatures will be, and knows it at signature generation time. This flexibility
on the opening values also leads to the
need of using extraction-based techniques for the modelling, since we cannot
expect that opening outputs the real identity of the signer. This comes as no
surprise, though, as it was already observed in GS schemes with non-conventional
opening \cite{dl21}, in DAA schemes \cite{cdl16,cdl16b} or in AC schemes that
need to ensure correctness of information generated before authentication, like
delegatable credentials \cite{bcc+09}, that prove validity of the delegation
chain. In fact, simulation-extractability seems to be most useful here,
something that was also observed in \cite{cl06} for producing signatures out of
knowledge of witnesses for concrete NP statements.

@@

\paragraph{Sample real-world use cases.} %
By enabling customization at issuance, signing and opening times, the
possibilities are large. As an example, consider a social network that requires
under-age users to get permission from one parent. This can be defined via a
delegatable credential scheme, checking that the delegation chain of the user
requesting to join is valid via a policy at issuance time. Later, in order to
make some post in the system advertising something to be sold in a local
second-hand page in the social network, the user may need to reveal, at signing
time, being a citizen of the local area (e.g. city). This can be done via an
appropriate policy at signing time. If, after the product is sold, it proves
to be defective, the social network may directly output the email of the user
if s/he is over-age, or the email of the adult in charge if s/he is under-age.
And that could be achieved via a policy for utility after signing time (at
opening time). Finally, if, in the same social network, some other
issuer-user-verifier situation comes up that requires a different set of utility
choices, just the appropriate functions would need to be changed, without the
need to re-model from scratch (or at all) -- and, given that our model allows
multiple issuers and openers, with different issuance and opening policies,
both could coexist.

%% Our techniques
%%% E.g., describe how we compatibilize both with AC sel disc., ring sigs, and MPS
\paragraph{Our construction approach.} %
The high-level idea for our generic construction \CUASGen is simple. Credentials
are signatures over blocks of messages (some signed in committed form), where
each message can be a different attribute, and a user who wants to
simultaneously leverage multiple credentials needs to bind all of them to the
same key. We add three generic functions: \fissue, leveraged by users at
credential issuance time to prove claims over attributes in other credentials,
in order to get a new one; \feval, proving correctness of the utility revealed
at signing time; and \finsp, proving correctness of the utility information
extractable after signing. The latter is, in addition, fed and modulated with
the output of \feval. The pattern of the statements proven in each function is
summarised in \figref{fig:proof-blocks-uas}.

@@

From \CUASGen, one can already build known (and more functionality-limited)
schemes from what we call \emph{restrictions} of \CUASGen simply by giving
concrete definitions of \fissue, \feval, and \finsp. Yet, the model remains the
same, and we prove that if such restrictions of \CUASGen are secure in the \UAS
model, they lead to secure constructions for the related primitives (GS
variants, AC variants, etc). To define our generic construction, we also specify
concrete NP relations for the underlying NIZK proofs for issuing, signing, and
opening. These NP relations must not be considered as immutable, though --
informally, it would seem that they must simply follow the pattern in
\figref{fig:proof-blocks-uas}. Actually, as mentioned, we show how by making a
minor variation into the relation used for signing, a \CUASGen restriction can
securely build ring signatures.

We emphasize that, while -- already before \UAS -- it may have not seemed hard
to build either of these primitives from another, the core of our contribution
is that, with \UAS, one essentially just needs to give concrete definitions of
three functions. Security of the final primitive then basically follows from
the fact that it is a secure instantiation of an \UAS scheme. We believe that
this is a powerful notion, that may finally allow research lines (AC and GS)
that have followed different paths to benefit from the findings of each other.
\figref{fig:relations} graphically summarises the relations we prove between
\UAS and other schemes -- although, of course, many more are probably
achievable.

\begin{figure}[ht!]
  \centering
  \scalebox{0.9}{
    \input{figures/relations.tex}
  }
  \caption{Summary of schemes that can be instantiated as restrictions of our
    generic constructions and proved secure under our \UAS model. Between
    parenthesis we show the adopted reference model, and the section offering
    full detail.}
  \label{fig:relations}
\end{figure}

### 2022-08-17 prism-backend/research/gsac/preliminaries-paper.tex
Migrating from fissue being a predicate to it being a function.
Applying changes after pass.

\paragraph{Digital Signatures.} %
Defined as a tuple (\SSetup,\SKeyGen,\SSign,\SVerify). Algorithm $\parm \gets
\SSetup(\secpar)$ produces public parameters for the other algorithms.
$(\vk,\sk) \gets \SKeyGen(\parm)$ generates the verification-signing key
pair, algorithm $\sig \gets \SSign(\sk,\msg)$ signs message \msg with
signing key \Ssk, producing signature \Ssig, and $1/0 \gets \SVerify(\vk,
\sig,\msg)$ checks whether \Ssig is a valid signature over \msg, under
verification key \Svk. We rely on existentially unforgeable digital signature
schemes.

@@

\paragraph{(Block) Commitment Schemes.} %
Defined as a tuple (\CSetup,\CCommit). Algorithm $\parm \gets
\CSetup(\secpar)$ produces the parameters for committing to values. $\Ccom
\gets \CCommit(\parm, \smsg; r)$ produces a commitment \Ccom over a set of
messages \smsg, using randomness $r$ (which we may omit when explicitly
stating it is not necessary). We will further use commitment schemes that
allow committing to vectors of messages. Therein, the \CCommit algorithm simply
receives a block of messages. In both cases, we require hiding and binding
(block) commitment schemes.

@@

We follow common notation habits. $y \gets A(x)$ means that algorithm $A$, on
input $x$, produces output $y$. When $A$ is a probabilistic process, and we want
to make explicit the randomness $r$, we use $y \gets A(x;r)$. $\langle y_A,y_B
\rangle \gets \langle A(x_A),B(x_B)\rangle$ denotes algorithms $A$ and $B$,
running an interactive protocol, where $A$ receives $x_A$ as input and produces
$y_A$ as output, and $B$ receives $x_B$ and produces $y_B$.
%
Our generic constructions rely on the previously defined building blocks, which
share common (but different)
tokens: e.g., we will resort to several NIZK proofs instances. To distinguish
them, we may use subscripts or superscripts: e.g., $\NIZKRel_A$ denotes the
NP relation used within some algorithm $A$, $\NIZKSetupRel$ means that
the setup algorithm of an underlying NIZK system is performed associated to
relation $\NIZKRel$; $\SBCMparm$ are the public parameters for an \SBCM scheme,
and $\Eparm$ are the public parameters for an encryption scheme, etc. Variables
written in bold font denote sets, e.g., while we use \ipk to denote a single
issuer public key, \sipk denotes a set of issuer's public keys. To refer to
a specific element of a set, we use sub-indexes, e.g., $\sipk_i$ is the $i$-th
\ipk within \sipk.
%
We also assume the existence of a well defined attribute space over which
issuers in our schemes can issue credentials on. We denote this set with
\AttrSpace.
%
Except when otherwise stated, we assume that, given a credential \cred, the
public key of its issuer, $\ipk_{\cred}$ can be retrieved. Similarly, the
attribute set contained in a credential \cred is denoted $\attrs_{\cred}$.

>>>>

% \paragraph{Digital Signatures.} %
% Defined as a tuple (\SSetup,\SKeyGen,\SSign,\SVerify). Algorithm $\parm \gets
% \SSetup(\secpar)$ produces public parameters for the other algorithms.
% $(\vk,\sk) \gets \SKeyGen(\parm)$ generates the verification-signing key
% pair, algorithm $\sig \gets \SSign(\sk,\msg)$ signs message \msg with
% signing key \Ssk, producing signature \Ssig, and $1/0 \gets \SVerify(\vk,
% \sig,\msg)$ checks whether \Ssig is a valid signature over \msg, under
% verification key \Svk. We rely on existentially unforgeable digital signature
% schemes.

@@

% \paragraph{(Block) Commitment Schemes.} %
% Defined as a tuple (\CSetup,\CCommit). Algorithm $\parm \gets
% \CSetup(\secpar)$ produces the parameters for committing to values. $\Ccom
% \gets \CCommit(\parm, \smsg; r)$ produces a commitment \Ccom over a set of
% messages \smsg, using randomness $r$ (which we may omit when explicitly
% stating it is not necessary). We will further use commitment schemes that
% allow committing to vectors of messages. Therein, the \CCommit algorithm simply
% receives a block of messages. In both cases, we require hiding and binding
% (block) commitment schemes.

@@

We follow common notation habits. $y \gets A(x)$ means that algorithm $A$, on
input $x$, produces output $y$. When $A$ is a probabilistic process, and we want
to make explicit the randomness $r$, we use $y \gets A(x;r)$. $\langle y_A,y_B
\rangle \gets \langle A(x_A),B(x_B)\rangle$ denotes algorithms $A$ and $B$,
running an interactive protocol, where $A$ receives $x_A$ as input and produces
$y_A$ as output, and $B$ receives $x_B$ and produces $y_B$.
%
Our generic constructions rely on the previously defined building blocks, which
share common (but different)
tokens: e.g., we will resort to several NIZK proofs instances. To distinguish
them, we may use subscripts or superscripts: e.g., $\NIZKRel_A$ denotes the
NP relation used within some algorithm $A$, $\NIZKSetupRel$ means that
the setup algorithm of an underlying NIZK system is performed associated to
relation $\NIZKRel$; $\SBCMparm$ are the public parameters for an \SBCM scheme,
and $\Eparm$ are the public parameters for an encryption scheme, etc. Variables
written in bold font denote sets, e.g., while we use \ipk to denote a single
issuer public key, \sipk denotes a set of issuer's public keys. To refer to
a specific element of a set, we use sub-indexes, e.g., $\sipk_i$ is the $i$-th
\ipk within \sipk.
%
We also assume the existence of a well defined attribute space over which
issuers in our schemes can issue credentials on. We denote this set with
\AttrSpace.
%
$0(p)$ returns $True$ if $p=0$, or $False$ otherwise.
